### 2005-03-14 Documentation/Manual.tex
PH -- Added pdf support. make pdf produces a file Manual.pdf
PH -- moving some files
Fig. earthint added - for some reason disappeared [WW]
*** empty log message ***
PH prepared the manual for the release of 2.0.11 and removed part 3 (modules) again since it becomes pertinent only for version 2.3

\sloppy
% \frenchspacing
\setlength{\unitlength}{1cm}

\usepackage{verbatim}
\usepackage{cite}
\usepackage[dvips]{color}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{a4wide}
\usepackage{fancyheadings}
\usepackage[footnotesize,bf]{caption}
\usepackage{makeidx}
%Preparations for multiple indices
\usepackage{multind}

@@

\begin{center}
\colorbox{black}{\includegraphics[width=7cm]{earthint}}

@@

\begin{center}
Version from \today\ for \GLOBES\ 2.0
\end{center}

@@

{\em\small $^{\mathrm{a},\mathrm{b}}$%
       Institut f\"ur Theoretische Physik, Physik--Department,\\
       Technische Universit\"at M\"unchen,
       James--Franck--Strasse, D--85748 Garching, Germany}

@@

\clearpage
\thispagestyle{empty}
\bigskip
\begin{quote}
    Copyright \copyright  2004  The GLoBES Team.
    Permission is granted to copy, distribute and/or modify this document
    under the terms of the GNU Free Documentation License, Version 1.2
    or any later version published by the Free Software Foundation;
    with the invariant Sections ``Terms of usage of \GLOBES" 
    and ``Acknowledgments'', no Front-Cover Texts, and no Back-Cover Texts.
    A copy of the license is included in the section entitled "GNU
    Free Documentation License".
\end{quote}
\bigskip

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PART III: Reference manual
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \input{part3}

@@

This work and the development of \GLOBES\ have been supported by:
\begin{itemize}
\item Physik Department der Technischen Universit\"at M\"unchen
\item Max-Planck-Institut f\"ur Physik 
\item Sonderforschungsbereich 375 f\"ur Astro-Teilchenphysik\\ 
der Deutschen Forschungsgemeinschaft
\item Studienstiftung des Deutschen Volkes 
\end{itemize}

>>>>

\sloppy
% \frenchspacing
\setlength{\unitlength}{1cm}
% Do not change following line pdfmanual depends on it!
\input{pdfswitch}
%\usepackage{pdfcolmk}
\usepackage{verbatim}
\usepackage{cite}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{a4wide}
\usepackage{fancyheadings}
\usepackage[footnotesize,bf]{caption}
\usepackage{makeidx}
%Preparations for multiple indices
\usepackage{multind}

@@

\begin{center}
\colorbox{black}{\includegraphics[width=7cm]{.save/earthint}}

@@

\begin{center}
Version from \today\ for \GLOBES\ 2.0.11
\end{center}

@@

{\em\small $^{\mathrm{a}}$%
       Physics Department,
       University of Wisconsin,
       1150 University Av., Madison, WI 53706, USA}

{\em\small $^{\mathrm{b}}$%
       Institut f\"ur Theoretische Physik, Physik--Department,\\
       Technische Universit\"at M\"unchen,
       James--Franck--Strasse, D--85748 Garching, Germany}

@@

\clearpage
\thispagestyle{empty}
\bigskip
\begin{quote}
    Copyright \copyright  2004,2005  The GLoBES Team.
    Permission is granted to copy, distribute and/or modify this document
    under the terms of the GNU Free Documentation License, Version 1.2
    or any later version published by the Free Software Foundation;
    with the invariant Sections ``Terms of usage of \GLOBES" 
    and ``Acknowledgments'', no Front-Cover Texts, and no Back-Cover Texts.
    A copy of the license is included in the section entitled "GNU
    Free Documentation License".
\end{quote}
\bigskip

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PART III: Module Support
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% not yet!
%\input{part3}

@@

This work and the development of \GLOBES\ have been supported by:
\begin{itemize}
\item Physik Department der Technischen Universit\"at M\"unchen
\item Max-Planck-Institut f\"ur Physik
\item Sonderforschungsbereich 375 f\"ur Astro-Teilchenphysik\\
der Deutschen Forschungsgemeinschaft
\item Studienstiftung des Deutschen Volkes
\item Institute for Advanced Study, Princeton
\item Physics Department of the University of Wisconsin, Madison
\end{itemize}

### 2005-03-14 Documentation/appendix.tex
last changes am Manual
PH prepared the manual for the release of 2.0.11 and removed part 3 (modules) again since it becomes pertinent only for version 2.3

The installation of \GLOBES\ is highly automated and there should not be
any problems on a decently up-to-date GNU/Linux system. The installation
has, however, only been tested on a limited number of platforms, mainly
running with various versions of SuSE Linux. Thus we would appreciate
to know your experiences with the installation on different platforms.
Please send an e-mail to <\bugs>.

@@

Besides the usual things such as a working libc, you need to have
\begin{itemize}
\item[gcc] The GNU compiler collection\\
\verb+gcc.gnu.org+
\item[BLAS] Basic Linear Algebra Subprograms\\
\verb+www.netlib.org/blas/+    
\item[LAPACK] Linear Algebra PACKage\\
\verb+www.netlib.org/lapack/+
\item[f2c] Fortran to C\\ 
\verb+www.netlib.org/f2c/+
\item[GSL] The GNU Scientific Library\\
\verb+www.gnu.org/software/gsl/+
\end{itemize}
The library \verb+libglobes+ should in principle compile with any
ANSI C/C++ compiler, but the \verb+globes+ binary uses the 
\verb+argp+ facility of \verb+glibc+ to parse its command line options.

All those libraries are also available as rpm's from the various
distributors of GNU/Linux: See their web sites for downloads. Due to
the rather difficult nature of the build process of BLAS and LAPACK,
it is recommended that one uses a suitable rpm instead. In addition,
there is a good chance that
gcc, f2c and GSL are already part of your installation. Furthmore, you need
a working \verb+make+ to build and install \GLOBES.

In future releases, it is planned to at least make LAPACK obsolete, because
then also BLAS and f2c would drop out. Furthemore, there is no priniciple
reason why one should not get rid of any C++ parts. 
This will greatly simplify
the build process and reduce the requirements for installation.

@@

The install target also will install a program with name \verb+globes+ to
\verb+$prefix/bin+.

If you are not using \verb+make install+ you will find the static libary
at \verb+source/.lib/libglobes.a+ which you can copy to any destination.
However keep in mind that the linking command will be somewhat different,
\ie\ you have to specify all the dynamically linked objects besides libglobes,
which are
\begin{quote}
{\tt
         -lm -lgsl -lgslcblas -lf2c -llapack -lblas
}
\end{quote}

In general it is advisable to use the shared libraries. If you don't
have root privileges, see the corresponding section 
(page~\pageref{inst_noroot}).
If you install \GLOBES\ with root privileges, do not forget to run 
{\tt ldconfig} after installation.

@@

\section*{Advanced topics}

@@

\input{gpl}

\input{fdl}

@@

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: Manual.tex
%%% End:

>>>>

The installation of \GLOBES\ is highly automated and there should not be
any problems on a decently up-to-date GNU/Linux system. The installation
has, however, only been tested on a limited number of platforms. The 
current version has been tested for SuSE, Red Hat, Fedora Core and
a 64bit version of SuSE. It should also compile and work on Mac OS. We 
would appreciate to know your experiences with the installation on different platforms.
Please send an e-mail to <\bugs>.

@@

Besides the usual things such as a working \verb`libc`, you need to have
\begin{itemize}
\item[gcc] The GNU compiler collection\\
\verb+gcc.gnu.org+
\item[GSL] The GNU Scientific Library\\
\verb+www.gnu.org/software/gsl/+
\end{itemize}
The library \verb+libglobes+ should in principle compile with any
ANSI C/C++ compiler, but the \verb+globes+ binary uses the 
\verb+argp+ facility of \verb+glibc+ to parse its command line options.
However, on platforms where \verb+argp+ is lacking \GLOBES\ has replacement code, 
thus it should also work there.

All those libraries are also available as rpm's from the various
distributors of GNU/Linux, see their web sites for downloads. Chances are 
that \verb+gcc+ and GSL are already part of your installation.  For building \GLOBES\ 
from source, however, not only working libaries for above packages are 
needed but also the headers,  especially for GSL. Depending on your
installation, eg. on RedHat/Fedora, of GSL this may require to additionally 
install a rpm-package named \verb+gsl-devel+. If GSL has been installed from the 
tar-ball as provided by \verb`gnu.org` no problems should occur. Furthmore you need 
a working \verb+make+ to build and install \GLOBES.

@@

The install target also will install a program with name \verb+globes+ to
\verb+$prefix/bin+ and the files in the \verb+data+ directory of the tar-ball
to \verb+$prefix/share/globes+.
\index{norm}{glb files@{\tt glb}-files!installation}

If you are not using \verb+make install+ you will find the static libary
at \verb+source/.lib/libglobes.a+ which you can copy to any destination.
However keep in mind that the linking command will be somewhat different,
\ie\ you have to specify all the dynamically linked objects besides libglobes.

The default install directory prefix is \verb+/usr/local+.  Consult the
"Further Information" section below for instructions on installing the
library in another location or changing other default compilation
options.

Moreover a config-script called \verb+globes-config+ will be installed. This
script displays all information necessary to link any program with GLoBES.
For building static libaries and linking against them see the corresponding
section of this file.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{LAPACK \& BLAS requirements}

GLoBES uses one single routine out of LAPACK (\verb^www.netlib.org/lapack/^) named
\verb^zgeev^ to diagonalize the oscillation Hamiltonian in matter. Since LAPACK
is not part of the standard installation of many distributions and we need
only one function, \GLOBES\ comes with a convenience version of the required
routines, which are taken from LAPACK and BLAS (\verb^www.netlib.org/blas^) and have
been translated to C with \verb^f2c^. In case your system has a working installation
of BLAS/LAPACK \GLOBES\ will use it. However, if your system is lacking BLAS/
LAPACK the convenience version will be compiled and linked into \GLOBES. If
for some reason BLAS/LAPACK is installed on the system and one still wishes
to use the convience version use \verb^--enable-lapack-convenience^. Using Fortran
code which has been translated to C with \verb^f2c^ requires a run-time library
called \verb^libf2c^ (on some systems it is called \verb^libg2c^). The configury of 
\GLOBES\ detects the presence of either of the two and takes care of it.
If none of the two is present, a convience library is used. If
for some reason \verb^libf2c^ is installed on the system and one still wishes
to use the convience version use \verb^--enable-libf2c-convenience^.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{GSL requirements}

Sometimes the GNU scientific library is not available or is installed in a non-standard
location. This situation can arise in an installation without root privileges. In this 
case one can specify \verb^--with-gsl-prefix=path_to_gsl^ as option to the \verb^configure^ script.
If one wants to use a shared version of \verb^libgsl^ then one has to make sure that the
linker finds the library at run-time. This can be achieved by setting the environment variable
\verb^LD_LIBRARY_PATH^ correctly, \ie\ (in bash)
\begin{quote} 
\verb^export LD_LIBRARY_PATH=path_to_gsl^
\end{quote}
You also can use a static version of GSL by either building \GLOBES\ with \verb^LDFLAG=-all-static^
or by configuring GSL with \verb^--disable-shared^. In both cases no further actions like setting
any environment variables is necessary.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Distributions}

\subsection*{Fedora Core 3}

FC3 is shipped with a non working version of LAPACK. Thus to
make \GLOBES\ work on FC3, configure \GLOBES\ with \verb'--enable-lapack-convenience', which
replaces the defunct LAPACK with \GLOBES' own version thereof. You can diagnose a 
non-working LAPACK by compiling one of the examples in the \verb^examples^ directory and 
trying to run it. It will hang forever in the function \GLB{SetRates()}.

\subsection*{RedHat -- all versions}

The standard rpm-based installation of GSL does not provide any header files for GSL, 
which are however needed to compile \GLOBES. You have to install an additional rpm-package
called \verb'gsl-devel'. Altnernatively you can install GSL from a tar-ball and use the 
verb'--with-gsl-prefix' option to the configure script of \GLOBES.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Advanced topics}

\subsection*{Building and Using static versions of \GLOBES}

In certain circumstances it may be useful to use a static version of libglobes
or any of the binaries, \eg\ when running on a cluster.

   The \verb+configure+ script accepts the option \verb+--disable-shared+, in which
case only static objects are built, i.e. only a static version of libglobes. 
In case your system does not support shared libraries the \verb+configure+ script 
recognizes this. If you give no options to \verb+configure+, both shared and static 
versions are built and will be installed. All binaries, however, will use dynamic 
linking. If you want to build static binaries, use \verb+LDFLAGS=-all-static+ for building
them.

   Sometimes it is convenient, eg. for debugging purposes, to have a statically
linked version of a program using \GLOBES\, which is easiest achieved by just linking
with \verb+libglobes.a+. If you need a completely statically linked version, please, have
a look at the Makefile in the \verb+examples+ directory.
\begin{quote}
\verb#make example-static#
\end{quote}
produces a statically linked program that should in principle run on most Linuxes.
It should be straightforward to adapt this example to your needs.

All these options rely on \verb+libtool+ and a working \verb`gcc` installation. It seems that \verb`gcc 3.x`
is broken in subtle way which makes it necessary to add a symbolic link in the \verb`gcc` library
directory. The diagnostic for this requirement is that building static programs fails with
the error message \verb+cannot find -lgcc_s+. In those cases, find \verb+libgcc.a+ and add a symbolic 
link in the same directory where you found it (this requires probably root privileges)
\begin{quote}
\verb#ln -s libgcc.a libgcc_s.a#
\end{quote}

If you can not write to this directory just use the following work around. Add the same link  
as above to the directory where you installed \GLOBES\ into
\begin{quote}
\verb#cd prefix/lib#
\verb#ln -s path_to_libgcc.a/libgcc.a libgcc_s.a#
\end{quote}
and then change back into the \verb+examples+ directory and type
\begin{quote}
\verb#make LDFLAGS=-Lprefix/lib example-static#
\end{quote}
and you are done.

If you don't have libtool on your system either, then you just can use the libtool script in
the \GLOBES\ tar-ball instead. Make sure it is in a place where it is found by \verb+make+.

@@

{\footnotesize
\input{gpl}

\input{fdl}
}

@@

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: Manual.tex
%%% TeX-master: "Manual"
%%% End:
